---
title: ""
author: ""
date: ""
output: 
  html_document:
    css: null
params:
  analysis_name: "Default Analysis"
  matrix_name: "Default Matrix"
  number_permut: "500"
  savepath: "Default Variable"
---


<style>
.flex-container {
  display: flex;
  align-items: flex-start;
}
.flex-image {
  flex: 1;
}
.flex-tables {
  flex: 2;
}

.centered-box {
  text-align: center;
  border: 2px solid black;
  padding: 10px;
  display: inline-block;
  width: 100%;
  box-sizing: border-box;
  margin-top: 20px;
  margin-bottom: 20px;
}

h1.title, h2.author, h3.date {
  margin: 0;
  padding: 0;
}
</style>

<div class="centered-box">
  <h1 class="title">VLAD sum-up of `r params$analysis_name` population in `r params$matrix_name` matrix</h1>
  <h3 class="date">Done the `r Sys.Date()`</h3>
</div>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE)
```



```{r load-data, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(dplyr)
library(tibble)
library(readr)
library(stringr)
library(ggplot2)
library(knitr)
library(htmltools)
library(png)
library(arrow)
library(ggstatsplot)
library(gtsummary)
library(rlang)

path_to_save <- params$savepath
number_permutations <-params$number_permut

main_matrix <- read_tsv(paste0(params$matrix_name, ".csv"))


# Read the file paths from the file
txt_files_paths <- readLines(paste0("To_analyse_with_R_", params$analysis_name, "_", params$matrix_name, ".txt"))
#variable_interest <- params$main_variable
variable_interest <- readLines(paste0("main_variable_", params$matrix_name, ".txt"), warn = FALSE)
variable_interest_len <- length(variable_interest)

#### On regarde quel type de variable c'est 
# Determine the type of analysis based on the unique values
if (variable_interest_len > 2) {
  type_analysis <- "anova"
} else if (variable_interest_len == 2) {
  type_analysis <- "group"
} else {
  type_analysis <- "continuous"
}

#Create a column for group if we are in ANOVA
if (type_analysis == "anova"){
  main_matrix$group <- Reduce(function(current_group, col_name) {
  ifelse(is.na(current_group) & main_matrix[[col_name]] == 1, col_name, current_group)
}, variable_interest, init = NA)
  main_matrix$group <- factor(main_matrix$group, levels = variable_interest)
} else if (type_analysis == "group") {
  # Add a new column 'group' to indicate the highest or lowest value
  main_matrix$group <- ifelse(main_matrix[[variable_interest[1]]] > main_matrix[[variable_interest[2]]],
                                variable_interest[1],
                                variable_interest[2])
  
  # Ensure 'group' is a factor with levels equal to the values in 'variable_interest'
  main_matrix$group <- factor(main_matrix$group, levels = variable_interest)
}


#### On en fait un tibble
combined_data <- tibble(main_matrix)

#### On fout toutes les valeurs au fur et à mesure dans le tableau
column_names <- c() ##Vecteur avec tous les noms des colonnes qu'on va rajouter

for (txt_file_path in txt_files_paths) {
  # Read the file without assuming a header
  mean_values <- read_tsv(txt_file_path, col_names = FALSE) %>%
    setNames(c("MeanValue"))  # Set the column name
  
  # Extract the parts of the filename without the directory and extension
  filename <- str_replace(basename(txt_file_path), "\\..*$", "")
  matches <- str_match(
    filename,
    "^meants__(.+?)__(.+?)__(.+?)__(.+?)__(.+?)__contrast(\\d+)$"
  )
  if (is.na(matches[1, 1])) {
    warning("Skipping file with unexpected name format: ", filename)
    next
  }
  met <- matches[1, 3]
  matrix_label <- matches[1, 4]
  acquisition_name <- matches[1, 5]
  mask_name <- matches[1, 6]
  contrast_name <- paste0("contrast", matches[1, 7])
  
  # Use parts of the filename to create a name for the column
  column_name <- paste(met, matrix_label, acquisition_name, mask_name, contrast_name, sep = "__")
  
  # Set the name of the column in the dataframe
  names(mean_values) <- column_name
  
  # Bind the data to the combined tibble
  combined_data <- bind_cols(combined_data, mean_values)
  
  # Append the new column name to the vector
  column_names <- c(column_names, column_name)
}


###====
#Saving the tibble into a feather database file

# Assuming params$analysis_name and params$matrix_name are defined
date <- Sys.Date()
unique_name <- paste0(params$analysis_name, "_", params$matrix_name, "_", date, ".feather")
file_path <- file.path(path_to_save, unique_name)

# Save the tibble to a new Feather file
write_feather(combined_data, file_path)

###==== 
#Image matrix
main_matrix_image_path <- paste0(params$matrix_name, "_visualisation.png")
main_matrix_correlation_path <- paste0(params$matrix_name, "_correlation_matrix.png")

```

## Sum-up of main matrix

<div class="flex-container">

<div class="flex-image">
<!-- HTML Image tag with dynamic src and title -->
<img src="`r main_matrix_image_path`" alt="Dynamic Image" width="500" />
</div>

<div class="flex-tables">

```{r sum_up_matrix, echo=FALSE, warning= FALSE, message= FALSE}

if (type_analysis != "continuous") {
table <- main_matrix %>% select(-any_of(variable_interest)) %>% tbl_summary(by = group, statistic = list(all_continuous() ~"{mean} ({sd})", all_categorical() ~ "{n} / {N} ({p}%)")) %>% add_p()
table
} else {
  table <- main_matrix %>% select(-Intercept) %>%tbl_summary(statistic = list(all_continuous() ~"{mean} ({sd})", all_categorical() ~ "{n} / {N} ({p}%)"))
  table
}
# Loop through the list of summary tables and print them
# for (colname in names(summary_tables)) {
#   cat(paste0("\n\n### Summary for ", colname, "\n"))
#   print(kable(summary_tables[[colname]], caption = paste("Summary Statistics of", colname)))
# }

```

<br />
<br />
<div style="text-align: center;">
<img src="`r main_matrix_correlation_path`" alt="Dynamic Image" width="300" />
</div>
</div>
</div>



## Sum-up of significative analyses

```{r sum_up_analyses, echo=FALSE, results='asis', include=TRUE}

# Looping through the columns and ploting them according to the type of analysis we are in

#print(combined_data)

for (column_name in column_names) {
  parts_column_name <- str_split(column_name, "__", simplify = TRUE)
  
  # Use parts of the filename to create a name for the column
  # Adjust the index numbers based on how you want to use the parts
  #column_name <- paste(parts[3], parts[4], parts[5], parts[6], sep = "_")
  Met <- parts_column_name[1]
  Matrix <- parts_column_name[2]
  acq <- parts_column_name[3]
  mask <- parts_column_name[4]
  contrast <- parts_column_name[5]
  
  if (type_analysis != "continuous") {
    plot <- ggbetweenstats(
  data = combined_data,
  x = group,
  y = !!sym(column_name),
  xlab = "Group",
  ylab = Met,
  bf.message = FALSE,
  caption = paste("File:", acq, ", Mask:", mask, ",", contrast, ", Nb permutations: ", number_permutations),
  pairwise.display = "significant",
  title = paste(Met,"by", Matrix, "in significative regions"),
  ggtheme = theme_classic(),  # You can use any theme you prefer
  point.args = list(size = 4, alpha = 0.8) 
) + 
  theme(
    axis.title.x = element_text(size = 16),  # Increase x axis label size
    axis.title.y = element_text(size = 16),  # Increase y axis label size
    axis.text.x = element_text(size = 14),   # Increase x axis text size
    plot.title = element_text(size = 20, hjust = 0.5)  # Center and increase title size
  )
    
  } else if (type_analysis == "continuous") {
    plot <- ggscatterstats(combined_data, 
                           x = !!sym(variable_interest), 
                           y = !!sym(column_name),
                           xlab = variable_interest, 
                           ylab = Met, 
                           bf.message = FALSE,
                           caption = paste("File:", acq, ", Mask:", mask, ",", contrast, ", Nb permutations: ", number_permutations),
                           ggtheme = theme_classic(),
                           point.args = list(size = 4, alpha = 0.8),
                           title = paste(Met,"by", Matrix, "in significative regions") )+ 
                             theme(
                               axis.title.x = element_text(size = 16),  # Increase x axis label size
                               axis.title.y = element_text(size = 16),  # Increase y axis label size
                               axis.text.x = element_text(size = 14),   # Increase x axis text size
                               plot.title = element_text(size = 20, hjust = 0.5)  # Center and increase title size
    )
  
  }
  
  
  
  # Define the title
  title <- paste("Visualization and plot of", Met, "based on", Matrix)
  
  # Display the section with a box around it
  cat('<div style="border: 2px solid black; padding: 10px; margin-bottom: 20px;">\n')
  
  # Print the title
  cat('<div style="text-align: center; font-size: 18px; font-weight: bold;">', title, '</div>')

  #print(plot)
  
  #print(column_name)
  name_plot <- paste0(path_to_save, "/", column_name, date, ".svg")
  ggsave(name_plot, plot = plot, width = 7.5, height = 6)
  
  # Create the path for the image
  image_path <- paste0(column_name, ".png")
  # Include the image using Markdown syntax
  cat('<div style="text-align: center;">\n')
  cat("\n\n![](", image_path, ")\n\n", sep = "")
  cat("\n\n![](", name_plot, ")\n\n", sep = "")
  cat('</div>')
  cat('</div>')
  
  
}

```
