---
title: ""
author: ""
date: ""
output:
  html_document:
    css: null
params:
  analysis_name: "Default Analysis"
  matrix_name: "Default Matrix"
  file_type: "mrsi"
  number_permut: "500"
  savepath: NULL
---

<style>
.flex-container {
  display: flex;
  align-items: flex-start;
}
.flex-image {
  flex: 1;
}
.flex-tables {
  flex: 2;
}

.centered-box {
  text-align: center;
  border: 2px solid black;
  padding: 10px;
  display: inline-block;
  width: 100%;
  box-sizing: border-box;
  margin-top: 20px;
  margin-bottom: 20px;
}

h1.title, h2.author, h3.date {
  margin: 0;
  padding: 0;
}
</style>

<div class="centered-box">
  <h1 class="title">SwE sum-up of `r params$analysis_name` population in `r params$matrix_name` matrix</h1>
  <h3 class="date">Done the `r Sys.Date()`</h3>
</div>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE)
```

```{r load-data, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(dplyr)
library(tibble)
library(readr)
library(stringr)
library(ggplot2)
library(svglite)
library(knitr)
library(htmltools)
library(arrow)
library(gtsummary)
library(lme4)
library(broom.mixed)
library(purrr)
library(tidyr)
library(rlang)
library(lmerTest)
library(MuMIn)
library(ggeffects)

`%||%` <- function(x, y) {
  if (is.null(x)) y else x
}

pick_existing <- function(paths) {
  for (p in paths) {
    if (file.exists(p)) return(p)
  }
  return(NULL)
}

read_matrix_block <- function(path) {
  lines <- readLines(path, warn = FALSE)
  start <- grep("^/Matrix", lines)
  if (length(start) > 0) {
    lines <- lines[(start + 1):length(lines)]
  }
  lines <- lines[nzchar(trimws(lines))]
  if (!length(lines)) return(NULL)
  suppressMessages(
    readr::read_table(
      paste(lines, collapse = "\n"),
      col_names = FALSE,
      show_col_types = FALSE
    )
  )
}

read_contrast_matrix <- function(matrix_name) {
  contrast_path <- pick_existing(c(
    paste0(matrix_name, ".con"),
    paste0(matrix_name, "_con.txt"),
    paste0(matrix_name, "_contrast.txt")
  ))
  if (is.null(contrast_path)) return(NULL)
  mat <- read_matrix_block(contrast_path)
  if (is.null(mat)) return(NULL)
  as.matrix(mat)
}

read_design_sub <- function(matrix_name) {
  sub_path <- pick_existing(c(
    paste0(matrix_name, "_design_sub.txt"),
    paste0(matrix_name, ".sub"),
    paste0(matrix_name, "-sub.sub")
  ))
  if (is.null(sub_path)) return(NULL)
  sub_df <- read_matrix_block(sub_path)
  if (is.null(sub_df)) return(NULL)
  colnames(sub_df) <- c("subject_id", "visit_id", "group_id")[seq_len(ncol(sub_df))]
  sub_df %>%
    mutate(across(everything(), as.factor))
}

guess_col <- function(df, patterns) {
  hits <- names(df)[str_detect(names(df), regex(paste(patterns, collapse = "|"), ignore_case = TRUE))]
  if (length(hits) > 0) hits[1] else NULL
}

format_terms <- function(terms) {
  vapply(terms, function(x) paste0("`", x, "`"), character(1))
}

max_cat_levels <- 6
force_categorical <- character(0)
force_continuous <- character(0)

is_integerish <- function(x, tol = 1e-8) {
  x <- x[!is.na(x)]
  if (!length(x)) return(FALSE)
  all(abs(x - round(x)) < tol)
}

is_binaryish <- function(x) {
  vals <- unique(na.omit(x))
  length(vals) == 2
}

is_categorical_var <- function(x, max_levels = max_cat_levels) {
  if (is.factor(x) || is.character(x) || is.logical(x)) return(TRUE)
  if (!is.numeric(x)) return(FALSE)
  vals <- unique(na.omit(x))
  if (!length(vals)) return(FALSE)
  (length(vals) <= max_levels && is_integerish(vals)) || is_binaryish(vals)
}

is_continuous_var <- function(x, max_levels = max_cat_levels) {
  is.numeric(x) && !is_categorical_var(x, max_levels = max_levels)
}

coerce_categorical <- function(df, cols, max_levels = max_cat_levels, force_cat = force_categorical, force_cont = force_continuous) {
  for (nm in cols) {
    if (!nm %in% names(df)) next
    if (nm %in% force_cont) next
    if (nm %in% force_cat || is_categorical_var(df[[nm]], max_levels = max_levels)) {
      df[[nm]] <- factor(df[[nm]])
    }
  }
  df
}

parse_interaction_parts <- function(term) {
  if (is.null(term)) return(NULL)
  if (grepl("_x_", term, fixed = TRUE)) return(strsplit(term, "_x_", fixed = TRUE)[[1]])
  if (grepl(":", term, fixed = TRUE)) return(strsplit(term, ":", fixed = TRUE)[[1]])
  if (grepl("*", term, fixed = TRUE)) return(strsplit(term, "*", fixed = TRUE)[[1]])
  NULL
}

path_to_save <- params$savepath %||% getwd()
number_permutations <- params$number_permut
matrix_basename <- params$matrix_name

matrix_csv_path <- paste0(matrix_basename, ".csv")
if (!file.exists(matrix_csv_path)) {
  stop("Matrix file ", matrix_csv_path, " not found. Cannot proceed.")
}

design_matrix <- read_tsv(matrix_csv_path, show_col_types = FALSE) %>%
  as_tibble()
design_terms <- names(design_matrix)

subject_col <- guess_col(design_matrix, c("subj", "subject", "id"))
visit_col <- guess_col(design_matrix, c("visit", "ses", "time", "tp"))
group_col <- guess_col(design_matrix, c("group", "grp"))

id_cols <- c(subject_col, visit_col) %>% discard(is.null)

design_terms_nuid <- design_terms[!design_terms %in% id_cols]
vars_for_type <- setdiff(design_terms_nuid, "Intercept")
design_matrix <- coerce_categorical(design_matrix, vars_for_type)

combined_data <- design_matrix

if (!is.null(group_col) && group_col %in% names(combined_data)) {
  combined_data[[group_col]] <- as.factor(combined_data[[group_col]])
}

txt_files_path <- paste0("To_analyse_with_R_", params$analysis_name, "_", params$matrix_name, ".txt")
if (!file.exists(txt_files_path)) {
  stop("File listing analyses for R (", txt_files_path, ") not found.")
}
txt_files_paths <- readLines(txt_files_path, warn = FALSE)

analysis_specs <- tibble()

for (txt_file_path in txt_files_paths) {
  if (!file.exists(txt_file_path)) next

  mean_values <- read_tsv(txt_file_path, col_names = FALSE, show_col_types = FALSE) %>%
    setNames(c("MeanValue"))

  filename <- str_replace(basename(txt_file_path), "\\..*$", "")
  matches <- str_match(
    filename,
    "^meants__(.+?)__(.+?)__(.+?)__(.+?)__(.+?)__contrast(\\d+)$"
  )
  if (is.na(matches[1, 1])) {
    warning("Skipping file with unexpected name format: ", filename)
    next
  }

  met <- matches[1, 3]
  matrix_label <- matches[1, 4]
  acquisition_name <- matches[1, 5]
  mask_name <- matches[1, 6]
  contrast_id <- suppressWarnings(as.integer(matches[1, 7]))
  contrast_id_val <- ifelse(is.na(contrast_id), 0L, contrast_id)

  base_name <- paste(met, matrix_label, acquisition_name, mask_name, sep = "__")
  column_name <- paste(base_name, paste0("contrast", contrast_id_val), sep = "__")

  if (nrow(mean_values) != nrow(combined_data)) {
    warning("Skipping ", txt_file_path, " because row counts do not match the matrix.")
    next
  }

  combined_data[[column_name]] <- mean_values[[1]]

  analysis_specs <- bind_rows(
    analysis_specs,
    tibble(
      column_name = column_name,
      contrast_id = contrast_id,
      base_name = base_name,
      brain_image = paste0(column_name, ".png")
    )
  )
}

if (nrow(analysis_specs) == 0) {
  stop("No valid analyses were loaded from ", txt_files_path)
}

# Save a feather backup of the merged data if a writable path is provided
if (!is.null(path_to_save)) {
  date_stamp <- Sys.Date()
  feather_name <- paste0(params$analysis_name, "_", matrix_basename, "_", date_stamp, "_swe.feather")
  feather_path <- file.path(path_to_save, feather_name)
  try(write_feather(combined_data, feather_path), silent = TRUE)
}

contrast_matrix <- read_contrast_matrix(matrix_basename)
contrast_terms <- design_terms_nuid
if (!is.null(contrast_matrix) && ncol(contrast_matrix) != length(contrast_terms)) {
  warning("Contrast file and design matrix have different numbers of columns. Aligning by the smallest dimension.")
  keep <- seq_len(min(ncol(contrast_matrix), length(contrast_terms)))
  contrast_matrix <- contrast_matrix[, keep, drop = FALSE]
  contrast_terms <- contrast_terms[keep]
}

main_matrix_image_path <- paste0(matrix_basename, "_visualisation.png")
main_matrix_correlation_path <- paste0(matrix_basename, "_correlation_matrix.png")
```

## Sum-up of main matrix

<div class="flex-container">

<div class="flex-image">
<!-- HTML Image tag with dynamic src and title -->
<img src="`r main_matrix_image_path`" alt="Design heatmap" width="500" />
</div>

<div class="flex-tables">

```{r sum_up_matrix, echo=FALSE, warning= FALSE, message= FALSE}
analysis_cols <- analysis_specs$column_name
summary_data <- design_matrix %>% select(-any_of(c(id_cols, analysis_cols)))

if (!is.null(group_col) && group_col %in% names(design_matrix)) {
  table <- design_matrix %>%
    select(-any_of(c(id_cols, analysis_cols))) %>%
    tbl_summary(
      by = all_of(group_col),
      statistic = list(all_continuous() ~ "{mean} ({sd})", all_categorical() ~ "{n} / {N} ({p}%)")
    ) %>%
    add_p()
} else {
  table <- summary_data %>%
    tbl_summary(
      statistic = list(all_continuous() ~ "{mean} ({sd})", all_categorical() ~ "{n} / {N} ({p}%)")
    )
}
table
```

<br />
<br />
<div style="text-align: center;">
<img src="`r main_matrix_correlation_path`" alt="Correlation heatmap" width="300" />
</div>
</div>
</div>

## Sum-up of significant analyses (linear mixed models)

```{r sum_up_analyses, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
analysis_cols <- analysis_specs$column_name
fixed_effects <- setdiff(design_terms_nuid, c("Intercept", id_cols))

for (idx in seq_len(nrow(analysis_specs))) {
  entry <- analysis_specs[idx, ]
  response <- entry$column_name
  contrast_id <- entry$contrast_id

  current_data <- combined_data %>%
    select(any_of(c(id_cols, design_terms, response))) %>%
    drop_na(all_of(response))

  if (!nrow(current_data)) next

  contrast_row <- NULL
  if (!is.null(contrast_matrix) && !is.na(contrast_id) &&
      contrast_id > 0 && contrast_id <= nrow(contrast_matrix)) {
    contrast_row <- contrast_matrix[contrast_id, , drop = TRUE]
  }

  main_predictors <- character(0)
  if (!is.null(contrast_row)) {
    main_predictors <- contrast_terms[which(abs(contrast_row) > 1e-6)]
    main_predictors <- setdiff(main_predictors, "Intercept")
  }

  x_var <- if (length(main_predictors) > 0) {
    main_predictors[1]
  } else if (!is.null(visit_col)) {
    visit_col
  } else if (length(fixed_effects) > 0) {
    fixed_effects[1]
  } else {
    NULL
  }

  # Build model formula
  rhs_terms <- setdiff(design_terms_nuid, c("Intercept", id_cols))
  safe_rhs <- format_terms(rhs_terms)
  rhs <- if (length(safe_rhs) > 0) paste(safe_rhs, collapse = " + ") else "1"
  if (!is.null(subject_col)) {
    rhs <- paste(rhs, paste0("(1 | `", subject_col, "`)"), sep = " + ")
  }
  form <- as.formula(paste0("`", response, "` ~ ", rhs))

  model <- tryCatch(
    withCallingHandlers(
      lmer(form, data = current_data, REML = FALSE),
      warning = function(w) {
        message("Model warning for ", response, ": ", conditionMessage(w))
        invokeRestart("muffleWarning")
      }
    ),
    error = function(e) NULL
  )

  if (is.null(model)) {
    next
  }
  


  fe_tab <- coef(summary(model))  # with lmerTest this contains df and p
  r2_vals <- tryCatch(r.squaredGLMM(model), error = function(e) NULL)

  beta <- se <- tval <- pval <- df_val <- ci_low <- ci_high <- NA_real_

  if (!is.null(x_var) && x_var %in% rownames(fe_tab)) {
    beta   <- fe_tab[x_var, "Estimate"]
    se     <- fe_tab[x_var, "Std. Error"]
    tval   <- fe_tab[x_var, "t value"]
    df_val <- if ("df" %in% colnames(fe_tab)) fe_tab[x_var, "df"] else NA_real_
    pval   <- if ("Pr(>|t|)" %in% colnames(fe_tab)) fe_tab[x_var, "Pr(>|t|)"] else NA_real_

    # 95% CI (you can switch method = "profile" if you like, but it's slower)
    ci <- suppressMessages(confint(model, parm = x_var, method = "Wald"))
    ci_low  <- ci[1]
    ci_high <- ci[2]
  }

  cat("\n**Fixed effect for** `", x_var, "`:\n\n", sep = "")
  if (!is.na(beta)) {
    cat(
      "- β = ", round(beta, 3),
      " [", round(ci_low, 3), ", ", round(ci_high, 3), "]",
      ", SE = ", round(se, 3),
      ", t", if (!is.na(df_val)) paste0("(", round(df_val, 1), ")") else "",
      " = ", round(tval, 2),
      if (!is.na(pval)) paste0(", p = ", signif(pval, 3)) else "",
      "\n", sep = ""
    )
  }

  if (!is.null(r2_vals)) {
    cat(
      "- R² (marginal) = ", round(r2_vals[1], 3),
      ", R² (conditional) = ", round(r2_vals[2], 3),
      "\n\n", sep = ""
    )
  }
  

  rhs_terms_no_ids <- setdiff(rhs_terms, "Intercept")

  title <- paste("Visualization and mixed-model of", entry$base_name, " - contrast", contrast_id)
  cat('<div style="border: 2px solid black; padding: 10px; margin-bottom: 20px;">\n')
  cat('<div style="text-align: center; font-size: 18px; font-weight: bold;">', title, '</div>')

  # Model display
  model_desc <- paste(response, "~", if (length(rhs_terms_no_ids) > 0) paste(rhs_terms_no_ids, collapse = " + ") else "1")
  if (!is.null(subject_col)) {
    model_desc <- paste(model_desc, "+ (1 |", subject_col, ")")
  }
  cat("\n\n**Model:** ", model_desc, "\n\n")

  plot_df <- current_data %>% drop_na(all_of(c(response, x_var)))

  # Default x variable for plotting and default color aesthetic
  plot_x_var <- x_var
  color_aes <- if (!is.null(group_col) && group_col %in% names(plot_df)) group_col else subject_col

  # Detect interaction terms encoded as Var1_x_Var2
  time_var  <- NULL
  group_var <- NULL
  interaction_parts <- parse_interaction_parts(x_var)
  is_interaction <- !is.null(interaction_parts)

  if (is_interaction) {
    parts <- interaction_parts
    if (length(parts) == 2 && all(parts %in% names(plot_df))) {
      var_a <- parts[1]
      var_b <- parts[2]
      is_a_cont <- is_continuous_var(plot_df[[var_a]])
      is_b_cont <- is_continuous_var(plot_df[[var_b]])
      is_a_cat  <- is_categorical_var(plot_df[[var_a]])
      is_b_cat  <- is_categorical_var(plot_df[[var_b]])

      # Only handle the case where one is continuous and the other is categorical
      if (is_a_cont && is_b_cat) {
        time_var  <- var_a
        group_var <- var_b
      } else if (is_b_cont && is_a_cat) {
        time_var  <- var_b
        group_var <- var_a
      }

      if (!is.null(time_var) && !is.null(group_var)) {

        # Ensure group variable is a factor
        if (!is.factor(plot_df[[group_var]])) {
          plot_df[[group_var]] <- factor(plot_df[[group_var]])
        }

        # For plotting, use the numeric variable on the x-axis and color by the group variable
        plot_x_var <- time_var
        color_aes  <- group_var
      }
    }
  }

  is_x_numeric <- !is.null(plot_x_var) && is_continuous_var(plot_df[[plot_x_var]])
  if (!is_x_numeric && !is.null(plot_x_var)) {
    plot_df[[plot_x_var]] <- factor(plot_df[[plot_x_var]])
  }
  if (!is.null(color_aes) && is_categorical_var(plot_df[[color_aes]])) {
    plot_df[[color_aes]] <- factor(plot_df[[color_aes]])
  }

  p <- ggplot(plot_df, aes(x = .data[[plot_x_var]], y = .data[[response]])) +
    theme_classic(base_size = 14) +
    labs(
      x = plot_x_var,
      y = response,
      title = paste0(entry$base_name, " (contrast ", contrast_id, ")"),
      caption = paste("Nb permutations:", number_permutations, "| File type:", params$file_type)
    )

  if (!is.null(subject_col) && is_x_numeric) {
    if (!is.null(color_aes)) {
      p <- p + geom_line(aes(group = .data[[subject_col]], color = .data[[color_aes]]), alpha = 0.35)
    } else {
      p <- p + geom_line(aes(group = .data[[subject_col]]), alpha = 0.35)
    }
  }
  if (!is.null(color_aes)) {
    if (is_x_numeric) {
      p <- p + geom_point(aes(color = .data[[color_aes]]), size = 2.5, alpha = 0.7)
    } else {
      p <- p + geom_jitter(aes(color = .data[[color_aes]]), width = 0.1, size = 2.5, alpha = 0.7)
    }
  } else {
    if (is_x_numeric) {
      p <- p + geom_point(size = 2.5, alpha = 0.7)
    } else {
      p <- p + geom_jitter(width = 0.1, size = 2.5, alpha = 0.7)
    }
  }

  if (is_interaction && !is.null(time_var) && !is.null(group_var)) {
    # NOTE:
    # The interaction in the design matrix is typically represented as a *product column*
    # (e.g. Age_W_x_Sex), not as a formula term Age_W:Sex.
    # ggpredict(model, terms=c(time_var, group_var)) does NOT automatically recompute
    # the product column when varying time_var and group_var, so it can yield nearly
    # identical slopes. We therefore build a prediction grid manually and recompute
    # any available "_x_" product terms.

    # Helper: robust numeric conversion for factor/character 0/1, 1/2, etc.
    as_num_safe <- function(v) {
      if (is.factor(v)) v <- as.character(v)
      out <- suppressWarnings(as.numeric(v))
      if (all(is.na(out))) {
        # fallback: map factor/character levels to 0..K-1
        out <- as.integer(factor(v)) - 1L
      }
      out
    }

    # Base grid: vary time_var across its observed range, for each level of group_var
    x_seq <- seq(
      min(plot_df[[time_var]], na.rm = TRUE),
      max(plot_df[[time_var]], na.rm = TRUE),
      length.out = 100
    )
    g_lev <- levels(factor(plot_df[[group_var]]))

    pred_g <- expand.grid(
      tmp_x = x_seq,
      tmp_g = g_lev,
      KEEP.OUT.ATTRS = FALSE,
      stringsAsFactors = FALSE
    )
    names(pred_g)[names(pred_g) == "tmp_x"] <- time_var
    names(pred_g)[names(pred_g) == "tmp_g"] <- group_var
    pred_g[[group_var]] <- factor(pred_g[[group_var]], levels = g_lev)

    # Fill other fixed effects with typical values (mean for numeric, reference for factor)
    for (nm in rhs_terms) {
      if (!nm %in% names(plot_df)) next
      if (nm %in% c(time_var, group_var)) next

      # If this is a product term like A_x_B and both components exist in pred_g,
      # recompute it instead of holding it constant.
      if (grepl("_x_", nm, fixed = TRUE)) {
        parts_nm <- strsplit(nm, "_x_", fixed = TRUE)[[1]]
        if (length(parts_nm) == 2 && all(parts_nm %in% names(pred_g))) {
          a <- parts_nm[1]; b <- parts_nm[2]
          pred_g[[nm]] <- as_num_safe(pred_g[[a]]) * as_num_safe(pred_g[[b]])
          next
        }
      }

      if (is.numeric(plot_df[[nm]])) {
        pred_g[[nm]] <- mean(plot_df[[nm]], na.rm = TRUE)
      } else {
        pred_g[[nm]] <- factor(plot_df[[nm]]) %>% levels() %>% .[1] %>% factor(levels = levels(factor(plot_df[[nm]])))
      }
    }

    # Also ensure the *tested* interaction column is recomputed if it is a product term
    if (!is.null(x_var) && grepl("_x_", x_var, fixed = TRUE) && x_var %in% rhs_terms) {
      parts_x <- strsplit(x_var, "_x_", fixed = TRUE)[[1]]
      if (length(parts_x) == 2 && all(parts_x %in% names(pred_g))) {
        pred_g[[x_var]] <- as_num_safe(pred_g[[parts_x[1]]]) * as_num_safe(pred_g[[parts_x[2]]])
      }
    }

    # Fixed-effects prediction (population-level)
    pred_g$predicted <- as.numeric(predict(model, newdata = pred_g, re.form = NA))

    p <- p +
      geom_line(
        data = pred_g,
        inherit.aes = FALSE,
        aes(x = .data[[time_var]], y = predicted, color = .data[[group_var]], group = .data[[group_var]]),
        linewidth = 1
      )

  } else if (is_x_numeric) {
    # Simple numeric predictor: one population-level slope
    pred_g <- ggpredict(model, terms = paste0(x_var, " [all]")) %>%
      as.data.frame()

    p <- p +
      geom_ribbon(
        data = pred_g,
        inherit.aes = FALSE,
        aes(x = x, ymin = conf.low, ymax = conf.high),
        alpha = 0.15
      ) +
      geom_line(
        data = pred_g,
        inherit.aes = FALSE,
        aes(x = x, y = predicted),
        linewidth = 1
      )

  } else if (!is.null(x_var)) {
    # Purely categorical predictor: show model-based means with CI
    pred_g <- ggpredict(model, terms = x_var) %>%
      as.data.frame()

    p <- p +
      geom_point(
        data = pred_g,
        inherit.aes = FALSE,
        aes(x = x, y = predicted),
        size = 3
      ) +
      geom_errorbar(
        data = pred_g,
        inherit.aes = FALSE,
        aes(x = x, ymin = conf.low, ymax = conf.high),
        width = 0.15
      )
  }

  subtitle_txt <- NULL
  if (!is.na(beta)) {
    subtitle_txt <- paste0(
      "β_", x_var, " = ", round(beta, 3),
      " [", round(ci_low, 3), ", ", round(ci_high, 3), "]",
      if (!is.na(pval)) paste0(", p = ", signif(pval, 2)) else "",
      if (!is.null(r2_vals)) paste0(
        " | R²m = ", round(r2_vals[1], 2),
        ", R²c = ", round(r2_vals[2], 2)
      ) else ""
    )
  }

  p <- p + labs(subtitle = subtitle_txt)

  # Save plot if possible
  plot_path <- file.path(path_to_save %||% getwd(), paste0(response, "_lme.svg"))
  try(
    ggsave(plot_path, plot = p, width = 8, height = 6, device = svglite::svglite),
    silent = TRUE
  )

  image_candidates <- c(
    paste0(entry$column_name, ".png"),
    paste0(entry$base_name, ".png")
  )
  image_path <- pick_existing(image_candidates) %||% image_candidates[1]

  cat('<div style="text-align: center;">\n')
  cat("\n\n![](", image_path, ")\n\n", sep = "")
  cat("\n\n![](", plot_path, ")\n\n", sep = "")
  cat('</div>')

  cat('</div>\n')
}
```
